---
title: "I <3 SF: Trees and Streets in my Favorite City <3 <3"
author: "Carbon Counters"
date: "1/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Attach Packages
library(tidyverse)
library(dplyr)
library(here)
library(sf)
library(tmap)
```

I am #blessed to get to look at SF trees, including Cole Valley trees, the best trees.

Below, I wrangle some data because Allison told me to. We'll find the top 5 highest observations of trees by legal status, wrangle, and graph. 
```{r}

# read in SF trees data

sf_trees <- read_csv(here::here("data", "sf_trees", "sf_trees.csv"))

# basic wrangling reminders

top_5_status <- sf_trees %>% 
  count(legal_status) %>% 
  drop_na(legal_status) %>% 
  rename(tree_count = n) %>% 
  relocate(tree_count) %>% 
  slice_max(tree_count, n = 5)

print(top_5_status)
```

And now, a 
#GRAPH

```{r}
ggplot(data = top_5_status, aes(x = fct_reorder(legal_status, tree_count), y = tree_count))+
  geom_col()+
  labs(x = "Legal Status", y = "Tree Count")+
  coord_flip()+
  theme_minimal()

```

How refreshing...Blackwood Acacia Trees. Finally, learning how to do text to columns in R

```{r}

blackwood <- sf_trees %>%
  filter(str_detect(species, "Blackwood Acacia")) %>% 
  select(legal_status, date, latitude, longitude)

ggplot(data = blackwood, aes(x = longitude, y = latitude))+
  geom_point()

```

little more wrangling - separate and unite (combine or separate columns)

```{r}

#separate
sf_trees_sep <- sf_trees %>% 
  separate(species, into = c("spp_scientific", "spp_common"), sep = "::")

#unite

sf_trees_unite <- sf_trees %>% 
  unite("id_status", tree_id:legal_status, sep = "LOVES")
```

Maps!!!! I'm gonna go rogue and not do what Allison did. We're going to first use `st_as_sf()` to convert lat and long to spatial coordinates.

```{r}
all_trees_spatial <- sf_trees_sep %>% 
  select(spp_common, latitude, longitude) %>%
  drop_na() %>% 
  st_as_sf(coords = c("longitude", "latitude"))

count_spp <- all_trees_spatial %>% 
  select(!geometry) %>% 
  count(spp_common) %>% 
  filter(spp_common != "") %>% 
  slice_max(n, n = 5)

top_tree <- unique(count_spp$spp_common)

all_trees_spatial <- all_trees_spatial %>% 
  filter(spp_common %in% c(top_tree))

st_crs(all_trees_spatial) = 4326

ggplot(data = all_trees_spatial)+
  geom_sf(aes(color = spp_common))
```

Read in SF roads shapefile
```{r}
sf_map <- read_sf(here("data", "sf_map", "tl_2017_06075_roads.shp"))

st_transform(sf_map, 4326)

ggplot(data = sf_map)+
  geom_sf()
```

Now we'll combine them!

```{r}
ggplot()+
  geom_sf(data = sf_map, size = 0.1, color = "darkgray")+
  geom_sf(data = all_trees_spatial, aes (color = spp_common), size = .1)+
  theme_void()
```

Let's make it interactive
```{r}
tmap_mode("view")

tm_shape(all_trees_spatial)+
  tm_dots()
```

